---
output: github_document
params:
  name: "your-name"
---

## Initialize pull request (PR)

* Initialize a PR with `pr_init("01-your-name")`.

* Copy this file as your-name.Rmd, e.g. mauro.Rmd.

## Global chunk options

* The first chunk of an .Rmd file is usually called `setup` and includes options that modify the output. These are my favourite options. What do they do? Experiment.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "#>",
  error = TRUE,
  collapse = TRUE
)
```

## Setup packages

* It's good practice to include all calls to `library()` at the top of your script. Why not include do this in the `setup` chunk? What does `include=FALSE` do in the `setup` chunk?

```{r}
library(praise)
```

## Chunk-specific options

* Make this chunk output comments of the form "x>". Why should chunk comments always start with `#`?

```{r praise}
praise()
```

* Commit with a meaningful message, e.g. "Tweak chunk options".

## Outputs

* Change `output:` to `pdf_document` and knit.

* Remove the .pdf output and revert the change to your-name.Rmd.

## Params

* In the YAML header above, edit `name: your-name` to prise yourself. How could you use `params` in real life?

```{r}
paste(params$name, praise())
```

## Submit

* Use `pr_push()` to submit a pull request with your changes to the .Rmd source and the .md output.

