---
title: "Rmarkdown"
output: github_document 
params:
  name: "Your name"
---

## Global chunk-options

* The first chunk of an .Rmd file is usually called `setup` and includes options that modify the output. These are my favourite options. What do they do? Experiment.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "#>",
  error = TRUE,
  collapse = TRUE
)
```

## Setup packages

* It's good practice to include all calls to `library()` at the top of your script. Why not include do this in the `setup` chunk? What does `include=FALSE` do in the `setup` chunk?

```{r}
library(praise)
```

## Chunk-specific options

* Make this chunk output comments of the form "x>". Why should chunk comments always start with `#`?

```{r praise}
praise()
```

* Commit with a meaningful message, e.g. "Tweak chunk options".

## Outputs

* Experiment with different outputs and knit each time:
  * `pdf_document`
  * `word_document`
  * `ioslides_presentation`.

* Revert to `github_document`, and remove all outputs except the .pdf output.

## Params

* Click on _Knit > Knit with Parameters ..._ and add your name so the code below prises you. How could you use `params` in real life?

```{r}
paste(params$name, praise())
```

