---
output: github_document
---

## Initialize pull request

* Remember to initialize a pull request with `pr_init()`.

* Rename this file with your name, e.g. from your-name.Rmd to mauro.Rmd.

* Commit with a meaningful message, e.g. "Personalise your-name.Rmd".

## Setup options and packages

* The first chunk of an .Rmd file is usually called "setup" and includes options, such as `echo = TRUE`, that modify the output. Kint this document; then add the options `comment = "#>"` and `collapse = TRUE` to `knitr::opts_chunk$set()`, and knit again. Based on your results, explain what these options do?:

  * `echo = TRUE`: ...
  * `collapse = TRUE`: ...
  * `comment = #>`: ...

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

* It's good practice to include all calls to `library()` at the top of your script. Why not include do this in the `setup` chunk?

```{r}
library(usethis)
```

* You can also pass options to specific chunks. Change the chunk `{r situation-report}` to `{r situation-report, comment = ">"}`. Why should chunk comments start with `#` instead of any other symbol (e.g. `>`)?

```{r situation-report}
usethis::git_sitrep()
```

* Commit with a meaningful message, e.g. "Tweak chunk options".

## Outputs

* Create a .pdf file: Change the `output:` to `pdf_document` and knit.

* Create a .word file: Change the `output:` to `word_document` and knit.

* Ignore all outputs except the .md file.

* Commit with a meaningful message, e.g. "Ignore .pdf and .docx outputs".

## Formatting

How is this chunk similar to the other chunks? How is it different?

```r
git_sitrep()
```

## Params

* TBD

## Submit

* Use `pr_push()` to submit a pull request with your changes to the .Rmd source and the .md output.


