---
title: Skateboard
output: github_document
---

**Collapse all chunks with Alt+O / Cmd+Option+O**



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, comment = "#>")
```



* Use the packages tidyverse and here.

```{r packages-1, eval = FALSE}
library(_________)
_______(here)
```

```{r packages-2, echo = FALSE}
library(tidyverse)
library(here)
```

You are about to read data but this code fails. The path is unsafe,  fix it:

* Create a safe path with `here()` and retry.
* Experiment with "Knit > Knit Directory".

```{r test-here, error=TRUE}
read_csv("data/gapminder.csv")
```



* Read the csv file; Create a safe path to "data/gampinder.csv" with `here()`.

```{r read-1, eval = FALSE}
gapminder <- ________(here("____", "gapminder.csv"))
```

```{r read-2, echo = FALSE}
gapminder <- read_csv(here("data", "gapminder.csv"))
```



* Print and explore `gampinder` under the chunk and with `view()`.

```{r gapminder-1, eval = FALSE}
_________
```

```{r gapminder-2, echo = FALSE}
gapminder
```



* Boxplot `lifeExp` for each continent `continent` (see `?geom_boxplot()`).

```{r boxplot-1, eval = FALSE}
ggplot(data = _________, aes(x = _________, y = _______)) +
  geom________()
```

```{r boxplot-2, echo = FALSE}
ggplot(data = gapminder, aes(x = continent, y = lifeExp)) +
  geom_boxplot()
```



## Takeaways

Import:

* Use `read_csv()` to read a single .csv file.
* Use `here()` to create safe paths with `here()`, particularly in rmarkdown.

Visualise:

* Start your plots from this template:

```
ggplot(data = <DATA>) +
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
```

Communicate: 

* Use github_document to report your analysis as a web page on GitHub.
* Set the chunk options you like.

Other:

* Use all packages in the tidyverse with `library(tidyverse)`.
* Assign values to objects with `<-`.
* Compose functions like `g(f(x))` or `x %>% f() %>% g()`.
