---
title: Skateboard
output: github_document
---

**Collapse all chunks with Alt+O / Cmd+Option+O**



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, comment = "#>")
```



* Use the packages tidyverse and here.

```{r packages-1, eval = FALSE}
library(_________)
_______(here)
```

```{r packages-2, echo = FALSE}
library(tidyverse)
library(here)
```



* Create a safe `path` to "data/gampinder" with `here()`.

```{r path-1, eval = FALSE}
path <- ____("data", "_________.csv")
```

```{r path-2, echo = FALSE}
path <- here("data", "gapminder.csv")
```



* Read the csv file and assign it to `gapminder`.

```{r read-1, eval = FALSE}
gapminder <- read____(____)
```

```{r read-2, echo = FALSE}
gapminder <- read_csv(path)
```



* Print and explore `gampinder` under the chunk and with `view()`.

```{r gapminder-1, eval = FALSE}
_________
```

```{r gapminder-2, echo = FALSE}
gapminder
```



* Boxplot `lifeExp` for each continent `continent` (see `?geom_boxplot()`).

```{r boxplot-1, eval = FALSE}
ggplot(data = _________, aes(x = _________, y = _______)) +
  geom________()
```

```{r boxplot-2, echo = FALSE}
ggplot(data = gapminder, aes(x = continent, y = lifeExp)) +
  geom_boxplot()
```



## Takeaways

* Use all packages in the tidyverse with `library(tidyverse)`.

Import:

* Use `read_csv()` to read a single .csv file.
* Use `here()` to craete safe paths to files -- particularly in rmarkdown.

Visualise:

* Start your plots from this template:

```
ggplot(data = <DATA>) +
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
```

Communicate: 

* Use github_document to report your analysis as a web page on GitHub.
* Set the chunk options you like.
