---
title: Scooter
output: github_document
params:
  continent:
    choices: [Africa, Americas, Asia, Europe, Oceania]
    multiple: true
    value: [Africa, Americas, Asia, Europe, Oceania]
---

**Collapse all chunks with Alt+O / Cmd+Option+O**



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, comment = "#>")
```

Packages.

```{r}
library(tidyverse)
library(here)
```

Import -- [practice safe paths](https://rstats.wtf/safe-paths.html).

```{r}
gapminder <- read_csv(here("data", "gapminder.csv"))
```



* Explore `gapminder` with `glimpse()`.

```{r glimpse-1, eval=FALSE}
gapminder %>% _______()
```

```{r glimpse-2, echo=FALSE}
gapminder %>% glimpse()
```



* Using `filter()`, subset rows where `continent` is "Europe".

```{r filter-continents-1, eval=FALSE}
continents <- c("________", "______")
_________ %>% ______(continent %in% continents)
```

```{r filter-continents-2, echo=FALSE}
continents <- "Europe"
gapminder %>% filter(continent %in% continents)
```



Notice the `value` of the `continent` `param` from the YAML header at the top.

```{r}
params
```



* Edit the YAML header so `param` `continent` takes the value Europe.
* Using `filter()`, subset rows where `continent` is `params$continent`.

```{r filter-params-continent-1, eval=FALSE}
filtered <- _________ %>% 
  ______(continent %in% ______$continent)

filtered
```

```{r filter-params-continent-2, echo=FALSE}
filtered <- gapminder %>% 
  filter(continent %in% params$continent)

filtered
```



* From `filtered`, exclude `country` and `continent` with `select()`.

```{r selected-1, eval=FALSE}
selected <- filtered %>% 
  ______(-country, -_________)
  select(-country, -continent)

selected
```

```{r selected-2, echo=FALSE}
selected <- filtered %>% 
  select(-country, -continent)

selected
```



Explore the relationship between `gdpPercap` and `lifeExp` with a scatterplot:

* Use `geom_point()`.
* As a third dimension, map point size to `pop`.
* Reduce overplotting by setting `alpha` to 1/3 and `shape` to `. Experiment.

```{r}
selected %>% 
  ggplot(aes(lifeExp, gdpPercap, size = pop)) + 
  geom_point(alpha = 1/3, shape = 1)
```



## Iteration

Above you used `select()` to exclude `country` and `continent`. Let's do this again but this time let's tackle the problem as an iteration challenge.

* Do it again but subset columns by type with `select(where(` and with `keep()`.
* Compare the result with `identical()`.

```{r iterate-by-type-1, eval=FALSE}
selected1 <- filtered %>% ______(where(is._______))
_________ %>% identical(selected)

selected2 <- filtered %>% ____(is._______)
_________ %>% identical(selected)
```

```{r iterate-by-type-2, echo=FALSE}
selected1 <- filtered %>% select(where(is.numeric))
selected1 %>% identical(selected)

selected2 <- filtered %>% keep(is.numeric)
selected2 %>% identical(selected)
```



Although it's not obvious, you've just solved an iteration problem. Let's make it explicit with `for()`:

```{r demo-for}
# Pre-allocate
pick <- logical(ncol(filtered))
# Iterate
for (i in seq_along(filtered)) {
  pick[[i]] <- is.numeric(filtered[[i]])
}
selected3 <- filtered[pick]
selected3 %>% identical(selected)
```

In this case, you could also discard character columns with `discard()`:

```{r demo-discard}
# Same
selected4 <- filtered %>% discard(is.character)
selected4 %>% identical(selected)
```

BTW, my iteration using `identical()` was inelegant; let's improve it:

```{r}
to_compare <- list(selected1, selected2, selected3, selected4)

to_compare %>% map_lgl(identical, selected)

# Same
to_compare %>% map_lgl(~identical(.x, selected))

# Same
to_compare %>% map_lgl(function(.x) identical(.x, selected))
```

You'll learn more about `map()` and friends later.



## Takeaways

Import:

* Use `glimpse()` to quickly explore the data you import.

Transform:

The most common and powerful data-science verbs are `filter()` and `select()`:

* Use with `filter()` to subset rows using column values.
* Use `select()` to subset columns using their names and types.

Visualise:

* Mapping and setting the same parameters produce different results.
* You can map data not only to `x`, `y` but also to `size`, `color`, etc.
* You may set `alpha`, `shape`, and `size` in ways that reduce overplotting.

Communicate: 

* You may use rmarkdown `params` to recreate reports with different parameters.

Iterate:

* You may avoid `for()` loops with `map()` & friends -- more on this later.
