---
title: Scooter
output: github_document
params:
  continent:
    choices: [Africa, Americas, Asia, Europe, Oceania]
    multiple: true
    value: [Africa, Americas, Asia, Europe, Oceania]
  year:
    value: 2017
    min: 1952
    max: 2007
    step: 1
    sep: ""
---

**Collapse all chunks with Alt+O / Cmd+Option+O**



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, comment = "#>")
```

* Packages.

```{r}
library(tidyverse)
library(here)
```

* Import -- [practice safe paths](https://rstats.wtf/safe-paths.html).

```{r}
gapminder <- read_csv(here("data", "gapminder.csv"))
```



# Your turn

* Explore.

```{r glimpse-3}
gapminder %>% 
  glimpse()
```

* Find `distinct()` values of `continent`.

```{r}
gapminder %>% distinct(continent)
```

* Subset rows where `continent` is "Americas" or "Europe".

```{r}
continents <- c("Americas", "Europe")
gapminder %>% filter(continent %in% continents)
```

* Subset rows where `continent` is one of `r params$continents`.

```{r}
continents <- c("Americas", "Europe")
gapminder %>% filter(continent %in% continents)
```


* Subset rows where `continent` is in `params$continents`.

```{r}
gapminder %>% filter(continent %in% params$continents)
```



Subset columns using their names and types:

* Select all columns except `country` and `continent`.

```{r}
gapminder %>% select()
```


```{r}
subset_cols <- gapminder %>% 
  select(continent, lifeExp, year) %>% 
  # Returns invisible.
  glimpse() %>% 
  filter(year >= 1980)

subset_cols
```





* Boxplot `lifeExp` for each continent `continent` (see `?geom_boxplot()`).

```{r boxplot}
ggplot(data = gapminder, aes(x = continent, y = lifeExp)) +
  geom_boxplot()
```
