---
title: "Bicycle"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, comment = "#>")
```



## Packages

```{r}
library(tidyverse)
library(here)
library(fs)
library(vroom)
```



## Definition of tidy data

<img src=http://i.imgur.com/SwZLZ3I.png width=750>



## From messy to tidy

<img src=http://i.imgur.com/ybegKuU.png width=750>



## Demo: A messy dataset

This toy dataset is messy. Why?

```{r demo-toy}
toy <- tribble(
  ~country, ~"2011", ~"2012", ~"2013",
      "fr",    7000,    6900,    7000,
      "de",    5800,    6000,    6200,
      "us",   15000,   14000,   13000
)
```



## Demo `pivot_longer()`

To tidy this dataset we pivot over all columns except `country`. We may exclude `country` with `-` or select all other columns with `where(is.numeric)`.

```{r}
toy %>% pivot_longer(where(is.numeric))
```



## Import

Import all datasets in "data/by-continent" into a single data frame: `messy`.

```{r}
paths <- dir_ls(here("data", "by-continent"))
messy <- suppressMessages(vroom(paths))
messy
```



## Tidy with `pivot_longer()`: Task

* Pivot over all numeric columns.
* Store the result as `longer`.
* Use `names_to = metric` (see `?pivot_longer()`).

```{r longer-1, eval=FALSE}
longer <- messy %>% 
  ____________(where(__________), ________ = "metric")

longer
```



## Tidy with `pivot_longer()`: Result

```{r longer-2, echo=FALSE}
longer <- messy %>% 
  pivot_longer(where(is.numeric), names_to = "metric")

longer
```



## `separate()`: Task

Mess things up again:

* `metric` is still messy; tidy it with `separate()` and `c("metric", "year")`.
* Store the result as `tidy`.

```{r separate-1, eval=FALSE}
tidy <- longer %>% ________(______, c("metric", "____"))
tidy
```



## `separate()`: Result

```{r separate-2, echo=FALSE}
tidy <- longer %>% separate(metric, c("metric", "year"))
tidy
```



## `unite()` and `pivot_wider()`: Task

* Use `unite()` to unite the columns `metric` and `year` as "metric".
* Use `pivot_wider()` to take the `names_from` the column `metric` and create new columns -- taking `values_from` the column `value`.
* Is the output wider than the input?

```{r mess-up-1, eval=FALSE}
tidy %>% 
  _____("metric", ______, year) %>% 
  ___________(names_from = ______, values_from = _____)
```



## `unite()` and `pivot_wider()`: Result

```{r mess-up-2, echo=FALSE}
tidy %>% 
  unite("metric", metric, year) %>% 
  pivot_wider(names_from = metric, values_from = value)
```



## Create a small dataset: Review

Let's create a small dataset to play with. Explain what this code does.

You should already understand this code:

- Remove `continent`.
- Subset life expectancy values for Argentina and Germany before 1962.
- Move the life expectancy values to the column `lifeExp`.



## Create a small dataset: Review

```{r}
subset1 <- tidy %>% 
  select(-continent) %>% 
  filter(
    country %in% c("Argentina", "Germany"), 
    metric == "lifeExp",
    year < 1962
  ) %>% 
  pivot_wider(names_from = metric)

subset1
```



## Create a small dataset: Task

Now add a new column `mean`, holding the mean `lifeExp` for each `country`:

* Use `group_by()`.
* Use `mutate()` to calculate mean `lifeExp` and to make `year` numeric.
* `ungroup()`.
* Store the result as `subset2`.

```{r group-mutate-1, eval=FALSE}
subset2 <- subset1 %>% 
  ________(country) %>% 
  ______(mean = ____(lifeExp), year = as.numeric(____)) %>% 
  _______()

subset2
```



## Create a small dataset: Result

```{r group-mutate-2, echo=FALSE}
subset2 <- subset1 %>% 
  group_by(country) %>% 
  mutate(mean = mean(lifeExp), year = as.numeric(year)) %>% 
  ungroup()

subset2
```



## Create a small dataset: Tweak

Let's degrade this dataset a bit for a later example.

```{r}
arg_ger <- subset2 %>% slice(-2) %>% select(-mean)
```



## `complete()`: What's missing?

Say you have the dataset `arg_ger`; it is missing data in between 1952 and 1957, but -- somehow -- you know the historical mean for each country. Let's use this data and knowledge to try fill the missing data.

```{r}
arg_ger

# Historical mean life expectancy
mean_argentina <- 63.5
mean_germany <- 68.3
```



## `complete()`: Task

* Make the implicit missing data explicit with `complete()`.

```{r complete-1, eval=FALSE}
arg_ger %>% ________(_______, year)
```



## `complete()`: Result

```{r complete-2, echo=FALSE}
arg_ger %>% complete(country, year)
```



## `fill`: Task

Extend the previous code:

* Fill the missing values of `lifeExp` with the value in `mean_argentina`.
* You'll need to pass each value to fill as a named list.
* Store the result as `filled`.

```{r fill-1, eval=FALSE}
filled <- arg_ger %>% 
  complete(country, year, fill = list(_______ = mean_argentina))

filled
```



## `fill`: Result

```{r fill-2, echo=FALSE}
filled <- arg_ger %>% 
  complete(country, year, fill = list(lifeExp = mean_argentina))

filled
```



## `full_seq()`: Task

Let's now complete the missing data in between 1952-1957.

* Use `pull()` to  pull the available years.
* Use `full_seq()` to produce the full sequence of every year in the range.
* Store the result as `all_years`.

```{r full-seq-1, eval=FALSE}
all_years <- filled %>% 
  ____(year) %>% 
  ________(period = 1)

all_years
```



## `full_seq()`: Result

```{r full-seq-2, echo=FALSE}
all_years <- filled %>% 
  pull(year) %>% 
  full_seq(period = 1)

all_years
```



## `fill()`: Task

* Use `all_years` to `complete()` `year`; also complete `country` with itself.
* Store the result as `full_mean`.

```{r all-years-1, eval=FALSE}
full_mean <- filled %>% ________(year = _________, country = _______)

full_mean
```



## `fill()`: Result

```{r all-years-2, echo=FALSE}
full_mean <- filled %>% complete(year = all_years, country = country)

full_mean
```



## `case_when()`: Task

Fill `lifeExp` with the mean historical values for each country:

* Use `mutate()` and `case_when()`.
* The 3 possible results are `mean_argentina`, `mean_germany`, or `lifeExp`.
* Store the result as `full`.



## `case_when()`: Task

```{r full-1, eval=FALSE}
full <- full_mean %>% 
  mutate(
    lifeExp = _________(
      _____(lifeExp) & country == "_________" ~ mean_argentina,
      is.na(_______) & _______ == "Germany"   ~ ____________,
      # Any other case
      TRUE                                    ~ lifeExp
    )
  )

full
```



## `case_when()`: Result

```{r full-2, echo=FALSE}
full <- full_mean %>% 
  mutate(
    lifeExp = case_when(
      is.na(lifeExp) & country == "Argentina" ~ mean_argentina,
      is.na(lifeExp) & country == "Germany"   ~ mean_germany,
      # Any other case
      TRUE                                    ~ lifeExp
    )
  )

full
```



## Plot: Task

* Make a line-plot of `year` versus `lifeExp`.
* Add a "dotted" line intercepting `y` at the historical mean of each country.
* Use `facet_wrap()` to plot each country in a separate panel.

```{r line-1, eval=FALSE}
full %>% 
  ggplot(aes(____, _______)) + 
  _____line() +
  _____hline(__________ = c(mean_argentina, mean_germany), linetype = "______") +
  __________(~country)
```



## Plot: Task

```{r line-2, echo=FALSE}
full %>% 
  ggplot(aes(year, lifeExp)) + 
  geom_line() +
  geom_hline(yintercept = c(mean_argentina, mean_germany), linetype = "dotted") +
  facet_wrap(~country)
```



## Create a presentation

* Run this code on the console to crate a presentation from this file.
* The output file is at "R/03_bicycle_you.html"; open it.

```
rmarkdown::render(
  input = here("R/03_bicycle_you.Rmd"), 
  output_format = "ioslides_presentation"
)
```



# Takeaways


## Takeaways: Imoprt

* Use `vroom()` to ready multiple files into a single data frame.



## Takeaways: Tidy

* The tidyverse prefers tidy data where:
  * Every variable is in its own column.
  * Every observation is in its own row.
* Use `pivot_longer()` and `pivot_wider()` to move columns to values and back.
* Use `unite()` to unite multiple ones into a single one.
* Use `separate()` to separate one column into multiple ones.
* Some functions you may use to fill missing values include:
  * `complete()`, `fill()`, `full_seq()`, `case_when()`.



## Takeaways: Transform

* The select-helpers (e.g. `where()`) help to transform but also to tidy data.
* You can `filter()` with multiple conditions separated by `&` or comma `,`.



## Takeaways: Visualise

* You can extend the basic plot template to include many layers.
* Use `facet_wrap()` to plot subset of data in separate panels.



## Takeaways: Communicate

You can transform your github_document into an .html slideshow with:

```
rmarkdown::render(
  input = here("your-github_document.Rmd"), 
  output_format = "ioslides_presentation"
)
```
