---
title: "Motorbike"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages.

```{r}
library(tidyverse)
```



## Warm up

This toy dataset is messy. Why?

```{r demo-toy}
toy <- tribble(
  ~country, ~"2011", ~"2012", ~"2013",
      "fr",    7000,    6900,    7000,
      "de",    5800,    6000,    6200,
      "us",   15000,   14000,   13000
)
```

To tidy this dataset we pivot over all columns except `country`. We may exclude `country` with `-` or select all other columns with `where(is.numeric)`.

```{r}
toy %>% pivot_longer(where(is.numeric))
```

This dataset is now tidy.



## Import

Import all datasets in "data/by-continent" into a single data frame: `messy`.

```{r}
paths <- dir_ls(here("data", "by-continent"))
paths

messy <- vroom(paths)
```



## Tidy

* Pivot over all numeric columns.
* Store the result as `longer`.
* Use `names_to = metric` (see `?pivot_longer()`).
* Compare the number of rows in the input and output.

```{r longer-1}

```
```{r longer-2, eval=FALSE}
longer <- messy %>% 
  ____________(where(__________), ________ = "metric")

longer
```

```{r longer-3, echo=FALSE}
longer <- messy %>% 
  pivot_longer(where(is.numeric), names_to = "metric")
longer
```



* `metric` is still messy; tidy it with `separate()` and `c("metric", "year")`.
* Store the result as `tidy`.

```{r separate-1}

```

```{r separate-2, eval=FALSE}
tidy <- longer %>% ________(______, c("metric", "____"))
```

```{r separate-3, echo=FALSE}
tidy <- longer %>% separate(metric, c("metric", "year"))
```




* Use `unite()` and `pivot_wider()` to mess things up back again.


TODO


## Filling missing data with `complete()`

This toy dataset has implicit missing data in year 2000. Why?

```{r}
kelp <- tibble(
  year = c(1999, 1999, 2000, 2004, 2004),
  taxon = c("Agarum", "Saccharina", "Saccharina", "Agarum", "Saccharina"),
  abundance = c(1, 4, 5, 8, 2)
)

kelp
```



* Use `complete()` to make explicit the missing data in `year` and `taxon`.

```{r complete-1}


```

```{r complete-2}
kelp %>% 
  ________(year, _____)

```

```{r complete-1}
kelp %>% 
  complete(year, taxon)
```



In `year = 2000` you found no individual of `taxon` "Agarum".

* Use the argument `fill` to fill missing data with the value `0`.

```{r fill-1}


```

```{r fill-2}
kelp %>% 
  complete(____, taxon, ____ = list(abundance = _))
```

```{r fill-3}
kelp %>% 
  complete(year, taxon, fill = list(abundance = 0))
```



You surveyed every year between 1999-2004 but found nothing between 2001-2003.

```{r}
success_years <- kelp %>% 
  distinct(year) %>% 
  pull(year)

success_years
```

* Use `full_seq()` to produce the full sequence of the years you surveyed.

```{r full-seq-1}




```

```{r full-seq-2}
all_years <- success_years %>% 
  ________(period = 1)

all_years
```

```{r full-seq-3}
all_years <- success_years %>% 
  full_seq(period = 1)

all_years
```



* Use `all_years` to `complete()` your dataset with explicit missing values.

```{r all-years-1}


```

```{r all-years-2}
kelp %>% 
  ________(year = _________, taxon = taxon)
```

```{r all-years-3}
kelp %>% 
  complete(year = all_years, taxon = taxon)
```



You now want to study "Agarum" only.

```{r}
agarum <- kelp %>% 
  filter(taxon == "Agarum") %>% 
  complete(year = all_years)

agarum
```



* `fill()` the missing values of `taxon`.

```{r fill-taxon-1}


```

```{r fill-taxon-2}
agarum ___ 
  ____(_____)
```

```{r fill-taxon-3}
agarum %>% 
  fill(taxon)
```

* Now also `replace_na()` in `abundance` with the value `0`.

```{r replace-na-1}



```

```{r replace-na-2}
agarum %>% 
  fill(taxon) %>% 
  __________(list(_________ = _))
```

```{r replace-na-3}
agarum %>% 
  fill(taxon) %>% 
  replace_na(list(abundance = 0))
```



***

# Take Aways

Data comes in many formats but the tidyverse prefers just one: _tidy data_.

A data set is tidy if and only if:

1. Every variable is in its own column.
2. Every observation is in its own row.
