---
title: "Bicycle"
output: github_document
---

**Collapse all chunks with Alt+O / Cmd+Option+O**



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, comment = "#>")
```

Packages.

```{r}
library(tidyverse)
library(here)
library(vroom)
library(fs)
library(glue)
```



This messy dataset comes in multiple files -- one per continent:

```{r demo-paths}
paths <- dir_ls(here("data", "by-continent"))
paths
```

You can use `map_df()` to interactively call `read_csv()` on each path.

```{r message=FALSE}
paths %>% map_df(read_csv)
```



But `vroom()` does this faster and in one step (it's "vectorized" over `file`).

* Read the data from all `paths` into a single data frame with `vroom()`.
* Pipe it into `glimpse()` and store the result as `messy` and print it.

```{r vroom-messy-2, eval=FALSE}
messy <- _____(paths) %>% _______()
messy
```

```{r vroom-messy-3, echo=FALSE}
messy <- vroom(paths) %>% glimpse()
messy
```



`glimpse()` is safe to use in pipelines. As it's called for its side effect, it returns its first argument `invisible()`.

```{r demo-invisible}
formalArgs(tibble:::glimpse.data.frame)

tail(tibble:::glimpse.data.frame)
```



A common way to explore data is by counting rows:

* `count()` rows by `continent`.
* Confirm no `country` has more than one row using `count()` and `filter()`.

```{r count-2, eval=FALSE}
messy %>% _____(continent)

messy %>% _____(country) %>% ______(_ > 1)
```

```{r count-3, echo=TRUE}
messy %>% count(continent)

messy %>% count(country) %>% filter(n > 1)
```



This messy dataset is hard to work with but not impossible. Let's study the mean life expectancy through time.

Let's focus on `continent` and all columns that `start_with()` "lifeExp":

```{r demo-starts-with}
life_exp <- messy %>% 
  select(continent, starts_with("lifeExp_"))

life_exp
```

The prefix "lifeExp" is redundant and we could remove it.

```{r demo-rename-with}
remove_prefix <- function(.x) str_replace(.x, "lifeExp_", "")

life_exp1 <- life_exp %>% 
  rename_with(remove_prefix)

# Same with the `~` shortcut, like in `map()` and friends
life_exp1 <- life_exp %>% 
  rename_with(~str_replace(.x, "lifeExp_", ""))

life_exp1
```

(The resulting names are non-syntactic so R surrounds them in backticks. Such names are awkward and you should avoid them. As you will soon see, here we won't refer to them directly; instead we'll use the tidyselect helpers -- see `?select()`.)



* Use `summarize()` to get the `mean` `across()` all columns except `continent`.

```{r summarize-across-2, eval=FALSE}
life_exp1 %>% 
  _________(across(-_________, mean))
```

```{r summarize-across-3, echo=FALSE}
life_exp1 %>% 
  summarize(across(-continent, mean))
```



* Now iterate over each `continent`, by grouping column-wise with `group_by()`.
* Compute across `everything()` (the grouping variable is excluded by default).
* Compute the number of observations per group (use `n = n()`).
* Relocate the `n` column to the front with `relocate()`.
* Store the result as `by_continent`.

```{r summarise-group-by-2, eval=FALSE}
by_continent <- life_exp1 %>% 
  ________(continent) %>% 
  summarize(across(everything(), mean), n = n()) %>% 
  ________(n)

by_continent
```

```{r summarise-group-by-3, echo=FALSE}
by_continent <- life_exp1 %>% 
  group_by(continent) %>% 
  summarize(across(everything(), mean), n = n()) %>% 
  relocate(n)

by_continent
```



* Now iterate over each year, by grouping row-wise with `rowwise()`.
* Create the new column `life_expectancy`: as `mean(c_across())`
* Create the new column `se`() standard error) as `sd(c_across()) / sqrt(n)`.

(See `?c_across()` -- its like `across()` but works with `rowwise()`.)

```{r summarize-rowwise-2, eval=FALSE}
total <- by_continent %>% 
  _______(continent) %>% 
  _________(life_expectancy = ____(c_across()), se = __(c_across()) / sqrt(_))

total
```

```{r summarize-rowwise-3, echo=FALSE}
total <- by_continent %>% 
  rowwise(continent, n) %>% 
  summarise(life_expectancy = mean(c_across()), se = sd(c_across()) / sqrt(n))

total
```



* Create a bar-plot (`?geom_col()`) of `continent` versus `life_expectancy`.
* Use `geom_errorbar()` to show uncertainty as `life_expectancy + se`.
* Add this title with `labs()`: "Mean life expectancy (1952-2007)".

```{r geom-col-2, eval=FALSE}
total %>% 
  ggplot(___(_________, life_expectancy)) +
  geom____() +
  _____________(aes(ymin = life_expectancy, ymax = _______________ + se)) +
  ____(_____ = "Mean life expectancy (1952-2007)")
```

```{r geom-col-3, echo=FALSE}
total %>% 
  ggplot(aes(continent, life_expectancy)) +
  geom_col() +
  geom_errorbar(aes(ymin = life_expectancy, ymax = life_expectancy + se)) +
  labs(title = "Mean life expectancy (1952-2007)")
```

That was hard work. R is column-oriented so it's best to first tidy the data.
