---
title: "Car"
output: github_document
params:
  year:
    value: [1952, 2007]
    min: 1952
    max: 2007
    step: 1
    sep: ""
---

**Collapse all chunks with Alt+O / Cmd+Option+O**



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, comment = "#>")
```



Packages.

```{r demo-packages}
library(tidyverse)
library(here)
library(vroom)
library(fs)
library(glue)
```



## Import

`vroom()` reads multiple datasets into a single data frame, elegantly and fast. But let's do it's job in a different way, breaking down what `vroom()` does into steps that help us understand the underlying iteration problem, and a solution we can latter apply to many other iteration problems.



Create a vector of paths to read .csv files from:

* Use `here()` to create a safe path to the folder "data/by-continent".
* Use `dir_ls()` to create a vector of paths to the .csv files in that folder.

```{r dir-ls-here-1, eval=FALSE}
paths <- ______(____("data", "by-continent"))
```

```{r dir-ls-here-2, echo=FALSE}
paths <- dir_ls(here("data", "by-continent"))
```



* `map()` each path to `read_csv` to create a list of data frames.
* `reduce()` the list of data frames to a single data frame with `bind_rows`.

```{r map-reduce-1, eval=FALSE, message=FALSE}
messy <- paths %>% 
  ___(________) %>%
  ______(bind_rows)

messy
```

```{r map-reduce-2, echo=FALSE, message=FALSE}
messy <- paths %>% 
  map(read_csv) %>%
  reduce(bind_rows)

messy
```



## Tidy

Create a longer dataset with a numeric column `year`, and one column per metric:

* Pivot over numeric columns.
* Separate the `metric` column into the new columns "metric" and "year".
* Mutate the `year` column to numeric.
* Pivot again, to create a wider data frame with each metric in its own column.

```{r tidy-1, eval=FALSE}
tidy <- messy %>% 
  ____________(_____(is.numeric), names_to = "metric") %>% 
  ________(metric, into = c("metric", "____")) %>% 
  ______(year = __________(year)) %>% 
  ___________(names_from = metric)

tidy
```

```{r tidy-2, echo=FALSE}
tidy <- messy %>% 
  pivot_longer(where(is.numeric), names_to = "metric") %>% 
  separate(metric, into = c("metric", "year")) %>% 
  mutate(year = as.numeric(year)) %>% 
  pivot_wider(names_from = metric)

tidy
```



## Transform

Subset the data you care about:

* Use `filter()` to pick years in the inclusive range of `params$year`.
* Use `select()` to pick `country`, `year`, and `lifeExp`.

```{r demo-params-year}
params$year
```

```{r pick-params-1, eval=FALSE}
picked <- tidy %>% 
  ______(year >= min(______$____) & year <= max(______$____)) %>% 
  ______(country, ____, lifeExp)

picked
```

```{r pick-params-2, echo=FALSE}
picked <- tidy %>% 
  filter(year >= min(params$year) & year <= max(params$year)) %>% 
  select(country, year, lifeExp)

picked
```



## Visualise

Plot a linear model of life expectancy through time:

* Map `group` to `country` (what happens if you don't).
* Use `geom_smooth()`; use the "lm" `method`.
* For clarity, remote the `se` shade, and reduce `alpha` and `size` to 1/3.

```{r plot-lines-1, eval=FALSE}
picked %>% 
  ggplot(aes(____, lifeExp, group = _______)) +
  ___________(method = "__", se = _____, alpha = ___, ____ = 1/3)
```

```{r plot-lines-2, echo=FALSE}
picked %>% 
  ggplot(aes(year, lifeExp, group = country)) +
  geom_smooth(method = "lm", se = FALSE, alpha = 1/3, size = 1/3)
```



## Model

Let's explore two ways to get and access the parameters of those models.

`map()` approach:

* `group_by()` `country` then `nest()`.
* Call `lm()` inside `map()` inside `mutate()`, mapping the list-column `data`.
* `map_*()` to apply `summary(.x)$r.squared` over the models list-column ...
* ... do this inside `summarise()`.

```{r map-approach-1, eval=FALSE}
rsq <- tidy %>% 
  ________(country) %>% 
  nest() %>% 
  mutate(mod = ___(data, ~lm(_______ ~ year, data = .x))) %>% 
  _________(rsq = map_dbl(mod, ~summary(.x)$r.squared))

rsq
```

```{r map-approach-2, echo=FALSE}
rsq <- tidy %>% 
  group_by(country) %>% 
  nest() %>% 
  mutate(mod = map(data, ~lm(lifeExp ~ year, data = .x))) %>% 
  summarise(rsq = map_dbl(mod, ~summary(.x)$r.squared))

rsq
```

Rowwise approach:

```{r rowwise-approach-1, eval=FALSE}
rsq2 <- tidy %>% 
  nest_by(country) %>% 
  ______(mod = list(__(_______ ~ year, data = data))) %>% 
  summarise(rsq = _______(mod)$r.squared)

rsq2
```

```{r rowwise-approach-2, echo=FALSE}
rsq2 <- tidy %>% 
  nest_by(country) %>% 
  mutate(mod = list(lm(lifeExp ~ year, data = data))) %>% 
  summarise(rsq = summary(mod)$r.squared)

rsq2
```



## Countries with highest and lowest `rsq`

In the same pipeline:

* `ungroup()` the dataset `rsq` (what happens if you don't?).
* `arrange()` to find order the data set by descending values of `rsq`
* `print()` to inspect the data at this point.
* `pull()` the `country` column.

```{r country-rsq-1, eval=FALSE}
countries <- rsq %>% 
  _______() %>%
  _______(____(rsq)) %>% 
  _____() %>% 
  ____(country)

head(countries)
```

```{r country-rsq-2, echo=FALSE}
countries <- rsq %>% 
  ungroup() %>%
  arrange(desc(rsq)) %>% 
  print() %>% 
  pull(country)

head(countries)
```



* `filter()` rows of `picked` where `country` is in the `head()` and `tail()` of `countries`.

```{r high-low-1, eval=FALSE}
high_low <- picked %>% 
  ______(country %in% c(____(countries), ____(countries)))

high_low
```

```{r high-low-2, echo=FALSE}
high_low <- picked %>% 
  filter(country %in% c(head(countries), tail(countries)))

high_low
```



## Do our models agree with those we get with ggplo2?

If so, our models with high (low) `rsq` should fit the data tightly (loosely).

Let's plot the countries with the 6 highest and 6 lowest values of `rsq`.

```{r}
make_plot <- function(data) {
  ggplot(data, aes(year, lifeExp)) + 
    geom_smooth(method = "lm") +
    facet_wrap(~country) +
    theme_bw()
}
```

We can use this helper function to plot all plots at once:

```{r}
high_low %>% make_plot()
```

We can also apply it to a list split by `country`, and save each plot to a file:

* `split()` `high_low` by `country`.
* `map()` to apply `make_plot`.
* Use `ggsave()` inside `iwalk()`, mapping the argument `plot` to `.x`., and the argument `filename` to paths in the "output" folder. Create such paths with `here()`, where the name of each .png files should come from mapping `.y` like so: `glue("{.y}.png")`.

```{r ggsave-iwalk-1, eval=FALSE}
high_low %>% 
  _____(.$country) %>% 
  map(_________) %>% 
  _____(~ ggsave(filename = here("output", glue("{__}.png")), plot = .x))
```

```{r ggsave-iwalk-2, echo=FALSE}
high_low %>% 
  split(.$country) %>% 
  map(make_plot) %>% 
  iwalk(~ ggsave(filename = here("output", glue("{.y}.png")), plot = .x))
```



* Knit with different parameters and inspect the output.



## Takeaways

Import:

* `vrooms()` is like `many_csv %>% map(read_csv) %>% reduce(bind_rows)`.

Transform:

* Use `nest_by()` to create a nested data frame with row-wise properties.
* Use `lm()` to fit a linear model to data.
  * With list-columns, use `map()` inside `mutate()`
  * With row-wise data, use `mutate()` or `summarise()` without `map()`.
* Use `pull()` to pull the values of a data frame column.
* Use `arrange()` and maybe `desc()` to reorder rows.
* Use `ungroup()` when you no longer need the grouping.

Tidy:

* Nest data frames: what happens in the data frame, stays in the data frame.

Visualise:

* Use `geom_smooth()` to plot linear models.
* Use `ggsave()` to save ggplots.

Communicate:

* You may pass multiple values to YAML prams with:

```yaml
params:
  value: [1952, 2007]
```

Iteration:

* Use `View()` to view lists (for data frames, prefer `dplyr::view()` ).
  * Use `reduce()` to reduce a list to a single value with a binary function.
  * Use the `walk*()` family to apply side-effect functions.
  * Use `iwalk()` as short hand for `map2(x, names(x), ...)`
