---
title: "Create data/ from data-raw/"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, comment = "#>")
```

This document creates the data in data/ from files in data-raw.

* Use tidyverse packages and friends.

```{r packages}
library(tidyverse)
library(here)
library(vroom)

library(fs)
library(glue)
```

Explore the root of this repository.

```{r}
# Call a function with no argument
here()

# Call a function with 1 argument
dir_ls(here())

# Same output, different syntax: nest
out1 <- dir_ls(here())

# Same output, different syntax: pipe
out2 <- here() %>% dir_ls()

# Call a function with 2 arguments
identical(out1, out2)

# Same
out1 %>% identical(out2)
```

Explore datasets.

```{r}
# Call a function with three arguments (one passed to ...) -- see `?dir_ls()`
here() %>% dir_ls(regex = "data", recurse = TRUE)
```

Import `wide` data from data-raw/.

```{r, message=FALSE}
wide <- vroom(here("data-raw", "gap_wide.csv"))

wide
```



## Beyond basics

This section shows `if()`, the not `!` operator, and the pattern `glue("{variable} string")`.

--



Find `continents`.

```{r}
continents <- wide %>% 
  distinct(continent) %>% 
  pull() %>% 
  tolower()

continents
```

Ensure data/by-continent/.

```{r}
by_continent_path <- here("data", "by-continent")

by_continent_path

if (!dir_exists(by_continent_path)) {
  dir_create(by_continent_path)
}
```

Create paths where to later save the data of each continent.

```{r}
paths <- path(by_continent_path, glue("{continents}.csv"))

paths
```



Split data by continent and save it in data/ (side effect with no output).

```{r}
wide %>% 
  group_split(continent) %>% 
  walk2(paths, write_csv)
```

Confirm.

```{r, message=FALSE}
# See files
dir_ls(by_continent_path)

# Read them back as a single data frame
df <- dir_ls(by_continent_path) %>% vroom()

df
```

Other purrr ways to do the same.

```{r, message=FALSE}
# Same
df1 <- dir_ls(by_continent_path) %>% 
  map(read_csv) %>% 
  reduce(bind_rows)

# Same
df2 <- dir_ls(by_continent_path) %>% 
  map_df(read_csv)

# Same
df3 <- dir_ls(by_continent_path) %>% 
  map_df(~read_csv(.x))

all_equal(df, df1)
all_equal(df, df2)
all_equal(df, df3)
```


