---
title: "Create files in data/ from files in data-raw/"
output: github_document
editor_options: 
  chunk_output_type: console
---

This document creates the data in data/ from files in data-raw. Also it helps introduce rmarkdown, R, tidyverse packages and friends, and the pipe.

## Skate

###  Introduction to rmarkdown, R, tidyverse packages and friends, and the pipe

rmarkdown: 

* Share reports online with rmarkdown's `output: github_document`.
* Write prose with Markdown and insert chunks of R code.
* Setup chunk.
* Set global chunk options.
* Set chunk-specific options.

R:

* Assign a value to an object.
* Call a function with 0, 1, 2, and 3 arguments.
* Passing arguments to `...`
* Compose functions by nesting.
* Compose functions with the pipe.
* Run code `if()` a condition is met.
* Not`!` operator (bang)

tidyverse:

* Use multiple packages at once.
* Notice conflicts.

Friends of the tidyverse:

* rmarkdown for communication.
* here to write safe paths (crucial in rmarkdown documents).
* fs to interact with the file system.
* vroom to read multiple files at once and fast.
* glue to paste strings in a readable way.

--



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, comment = "#>")
```

* Use tidyverse packages and friends.

```{r packages}
library(tidyverse)
library(here)
library(vroom)

library(fs)
library(glue)
```

Explore the root of this repository.

```{r}
# Call a function with no argument
here()

# Call a function with 1 argument
dir_ls(here())

# Same output, different syntax: nest
out1 <- dir_ls(here())

# Same output, different syntax: pipe
out2 <- here() %>% dir_ls()

# Call a function with 2 arguments
identical(out1, out2)

# Same
out1 %>% identical(out2)
```

Explore datasets.

```{r}
# Call a function with three arguments (one passed to ...) -- see `?dir_ls()`
here() %>% dir_ls(regex = "data", recurse = TRUE)
```

Import `wide` data from data-raw/.

```{r, message=FALSE}
wide <- vroom(here("data-raw", "gap_wide.csv"))

wide
```

Find `continents`.

```{r}
continents <- wide %>% 
  distinct(continent) %>% 
  pull() %>% 
  tolower()

continents
```

Ensure data/by-continent.

```{r}
by_continent_path <- here("data", "by-continent")
by_continent_path

if (!dir_exists(by_continent_path)) {
  dir_create(by_continent_path)
}

paths <- path(by_continent_path, glue("{continents}.csv"))
```



## Drive

### Introduction to iteration with purrr

* Split a data frame by groups.
* Apply a function to each element of a list or atomic vector
* Iterate over multiple inputs simultaneously.
* Apply a function called primarily for its side effect.
* Specify the expected output with specialized, `map_*()` functions.
* Inline a function with the formula shortcut `~`.

--



Split data by continent and save it in data/ (side effect with no output).

```{r}
wide %>% 
  group_split(continent) %>% 
  walk2(paths, write_csv)
```

Confirm.

```{r, message=FALSE}
# See files
dir_ls(by_continent_path)

df <- dir_ls(by_continent_path) %>% vroom()
df
# Read them back as a single data frame
df1 <- dir_ls(by_continent_path) %>% map(read_csv) %>% reduce(bind_rows)

# Same
df2 <- dir_ls(by_continent_path) %>% map_df(read_csv)

# Same
df3 <- map_df(.x = dir_ls(by_continent_path), ~read_csv(.x))
```

